---
title: "RE6013 Business Analytics"
output: html_notebook
---

# Loading Data and Imports

Get imports
```{r}
library(data.table)
library(ggplot2)
library(corrplot)
library(DMwR2)
```

Load data
```{r}
home.train.dt <- fread("home_train.csv")
home.test.dt <- fread("home_test.csv")
```

# Data Exploration

Create exploration copy
```{r}
home.exp <- copy(home.train.dt)
```

Basic data check reveal the following continous columns:
  - Continuous
    - ApplicantIncome
    - CoApplicantIncome
    - LoanAmount
    - Loan_Amount_Term
    
Credit_History is actually categorical
```{r}
summary(home.exp) #13 columns in pattern, including 1 label column
```

Preview of the values reveal tha Loan_Status is the label column, Categorical and labelled with 'Y' and 'N'
```{r}
head(home.exp)
```

Check distribution of non-continuous variables

We can see missing values in the following non-continuous columns:
- Gender: missing is ""
- Married: missing is ""
- Dependents: missing is ""
- Self_Employed: missing is ""
- Credit_History: missing is NA
```{r}
ggplot(home.exp, aes(Gender)) + geom_histogram(stat="count")
ggplot(home.exp, aes(Married)) + geom_histogram(stat="count")
ggplot(home.exp, aes(Dependents)) + geom_histogram(stat="count")
ggplot(home.exp, aes(Education)) + geom_histogram(stat="count")
ggplot(home.exp, aes(Self_Employed)) + geom_histogram(stat="count")
ggplot(home.exp, aes(Property_Area)) + geom_histogram(stat="count")
ggplot(home.exp, aes(Loan_Status)) + geom_histogram(stat="count")
ggplot(home.exp, aes(Credit_History)) + geom_histogram(stat="count")
```
Non-continuous variables missing counts
```{r}
home.exp[Gender == "", .N]
home.exp[Married == "", .N]
home.exp[Dependents == "", .N]
home.exp[Self_Employed == "", .N]
home.exp[is.na(Credit_History), .N]
```
Continuous variables missing counts
```{r}
home.exp[is.na(ApplicantIncome), .N]
home.exp[is.na(CoapplicantIncome), .N]
home.exp[is.na(LoanAmount), .N]
home.exp[is.na(Loan_Amount_Term), .N]
```
# Filling missing categorical data

The missing data maybe filled in the following ways:
- Gender, Dependents, Self_Employed, Credit_History: use mode to impute
- Married: small enough to try profiling and using mode
- LoanAmount and Loan_Amount_Term: may need to discard since it is what we are predicting

```{r}
# get mode function definition
getmode <- function(v) {
   uniqv <- unique(v)
   uniqv[which.max(tabulate(match(v, uniqv)))]
}
```

Create duplicate
```{r}
home <- copy(home.exp)
```

Impute with mode for Gender and Dependents column
```{r}
#get modes
gender.mode <- getmode(home.exp$Gender)
dependents.mode <- getmode(home.exp$Dependents)
self_employed.mode <- getmode(home.exp$Self_Employed)

#impute with modes
home[Gender=="", Gender:=gender.mode]
home[Dependents=="", Dependents:=dependents.mode]
home[Self_Employed=="", Self_Employed:=self_employed.mode]
home[is.na(Credit_History), Credit_History:=0] #reasonable to assume missing is none

#verify
levels(factor(home$Gender))
levels(factor(home$Dependents))
levels(factor(home$Self_Employed))
levels(factor(home$Credit_History))
```

Perform some profiling on the 3 missing married patterns
Use Dependents and CoapplicantIncome
```{r}
# view the 3 with married missing
home[Married==""]

# get modes
LP001357.married.mode = getmode(home[Dependents == 0 & CoapplicantIncome > 0]$Married)
LP001760.LP002393.married.mode = getmode(home[Dependents == 0 & CoapplicantIncome == 0]$Married)

# impute with modes
home[Loan_ID=="LP001357", Married:=LP001357.married.mode]
home[Loan_ID=="LP001760" | Loan_ID=="LP002393", Married:=LP001760.LP002393.married.mode]

# verify
home[Loan_ID=="LP001357" | Loan_ID=="LP001760" | Loan_ID=="LP002393"]
```
```{r}
summary(home) #verify
```

# View distributions on filled data

It can be seen that the continuous variables have a small interquartile range in comparison to outliers - hard to see any distribution
```{r}
ggplot(home, aes(ApplicantIncome)) + geom_boxplot()
ggplot(home, aes(CoapplicantIncome)) + geom_boxplot()
ggplot(home[!is.na(LoanAmount)], aes(LoanAmount)) + geom_boxplot()
ggplot(home[!is.na(Loan_Amount_Term)], aes(Loan_Amount_Term)) + geom_boxplot()
```
View distributions in their raw state
- Notice that Loan_Amount_Term is overwhelmingly 360 days (1 year)
```{r}
ggplot(home, aes(ApplicantIncome)) + geom_histogram()
ggplot(home, aes(CoapplicantIncome)) + geom_histogram()
ggplot(home[!is.na(LoanAmount)], aes(LoanAmount)) + geom_histogram()
ggplot(home[!is.na(Loan_Amount_Term)], aes(Loan_Amount_Term)) + geom_histogram()
```

View distribution of...
- Applicantincome, Coapplicantincome and LoanAmount without outliers
- LoanAmountTerm for non-360
```{r}
# grab range for non-outliers
ai.q <- quantile(home$ApplicantIncome, probs = c(.25, .75), na.rm = FALSE)
ai.iqr <- IQR(home$ApplicantIncome)
cai.q <- quantile(home$CoapplicantIncome, probs = c(.25, .75), na.rm = FALSE)
cai.iqr <- IQR(home$CoapplicantIncome)
la.q <- quantile(home[!is.na(LoanAmount)]$LoanAmount, probs = c(.25, .75), na.rm = FALSE)
la.iqr <- IQR(home[!is.na(LoanAmount)]$LoanAmount)

# plot
ggplot(home[ApplicantIncome > ai.q[1]-1.5*ai.iqr & ApplicantIncome < ai.q[2]+1.5*ai.iqr], aes(ApplicantIncome)) + geom_histogram(binwidth = 50)
ggplot(home[CoapplicantIncome > cai.q[1]-1.5*cai.iqr & CoapplicantIncome <= cai.q[2]+1.5*cai.iqr], aes(CoapplicantIncome)) + geom_histogram(binwidth = 50)
ggplot(home[!is.na(LoanAmount) & LoanAmount > la.q[1]-1.5*la.iqr & LoanAmount < la.q[2]-1.5*la.iqr], aes(LoanAmount)) + geom_histogram(binwidth = 1)
ggplot(home[!is.na(Loan_Amount_Term) & Loan_Amount_Term != 360], aes(Loan_Amount_Term)) + geom_histogram(binwidth = 10)
```
Since there is large amount of people with 0 Coapplicant, see distribution of non-zero, barring outliers
```{r}
ggplot(home[CoapplicantIncome > cai.q[1]-1.5*cai.iqr & CoapplicantIncome <= cai.q[2]+1.5*cai.iqr & CoapplicantIncome != 0], aes(CoapplicantIncome)) + geom_histogram(binwidth = 50)
```
# Data cleaning for Loan Amount to give
Find correlations for approved loans
- LoanAmount and ApplicantIncome seem to have a somewhat positive correlation
```{r}
corr.matrix <- cor(home[Loan_Status=="Y",c('ApplicantIncome','CoapplicantIncome','LoanAmount','Loan_Amount_Term')], use='complete.obs')
corrplot(corr.matrix, method="color",addCoef.col = "black",tl.col = "black")
```
Try looking at scatter plot
- it appears that the correlation may only be present because of outliers
```{r}
ggplot(home[Loan_Status=="Y"], aes(x=ApplicantIncome, y=LoanAmount)) + geom_point()
```
Try scatterplot and correlation without outliers
- correlation becomes less strong but still visible
```{r}
corr.matrix <- cor(home[Loan_Status=="Y" & ApplicantIncome < 20000,c('ApplicantIncome', 'LoanAmount')], use='complete.obs')
corrplot(corr.matrix, method="color",addCoef.col = "black",tl.col = "black")
ggplot(home[Loan_Status=="Y" & ApplicantIncome < 20000], aes(x=ApplicantIncome, y=LoanAmount)) + geom_point()
```

## Feature Engineering
Maybe create new features?
- impute missing values
- combined income
- LoanAmount / LoanAmountTerm * 30 to compare with income
```{r}
# copy
home.loan <- copy(home[Loan_Status=="Y"])

# impute rows with missing loan information using mean
loanamount.mean <- as.integer(mean(home.loan$LoanAmount, na.rm = TRUE))
loanamountterm.mean <- as.integer(mean(home.loan$Loan_Amount_Term, na.rm=TRUE))
home.loan[is.na(LoanAmount), LoanAmount:=loanamount.mean]
home.loan[is.na(Loan_Amount_Term), Loan_Amount_Term:=loanamountterm.mean]

# verify imputation
sum(is.na(home.loan$LoanAmount))
sum(is.na(home.loan$Loan_Amount_Term))

# combine income
home.loan[,Income:=as.integer(ApplicantIncome+CoapplicantIncome)]
# create new loan amount
home.loan[,LoanMonthly:=LoanAmount/Loan_Amount_Term*30]

# verify
head(home.loan)
```
Check new correlations
- random outlier (Loanmonthly > 250) is skewing correlations (should remove in cleaning)
```{r}
corr.matrix <- cor(home.loan[,c('Income', 'LoanMonthly')], use='complete.obs')
corrplot(corr.matrix, method="color",addCoef.col = "black",tl.col = "black")
ggplot(home.loan, aes(x=Income, y=LoanMonthly)) + geom_point()
```
Find correlations without outlier
- much stronger and better
```{r}
corr.matrix <- cor(home.loan[LoanMonthly<250,c('Income', 'LoanMonthly', 'LoanAmount', 'Loan_Amount_Term')], use='complete.obs')
corrplot(corr.matrix, method="color",addCoef.col = "black",tl.col = "black")
ggplot(home.loan[LoanMonthly<250], aes(x=Income, y=LoanMonthly)) + geom_point()
```
See distribution wihtout outlier
```{r}
ggplot(home.loan[LoanMonthly<250], aes(LoanMonthly)) + geom_histogram(binwidth = 5)
mean(home.loan$LoanMonthly)
```
Remove outlier
```{r}
home.loan.no.outlier <- home.loan[LoanMonthly < 250]
ggplot(home.loan.no.outlier, aes(x=Income, y=LoanMonthly)) + geom_point()
```
Factor the categorical columns
```{r}
cols <- c("Gender", "Married", "Dependents", "Education", "Self_Employed", "Property_Area", "Loan_Status")
home.loan.no.outlier[,cols] <- lapply(home.loan.no.outlier[,cols, with=FALSE], factor)
```

```{r}
summary(home.loan.no.outlier)
```

Save file
```{r}
fwrite(home.loan.no.outlier, file="home_loan_structure_prediction.csv")
```

# Data cleaning for Loan Eligibility
Perform imputation + feature engineering
```{r}
# copy
home.eligibility <- copy(home)

# impute rows with missing loan information using mean
loanamount.mean <- as.integer(mean(home.eligibility$LoanAmount, na.rm = TRUE))
loanamountterm.mean <- as.integer(mean(home.eligibility$Loan_Amount_Term, na.rm=TRUE))
home.eligibility[is.na(LoanAmount), LoanAmount:=loanamount.mean]
home.eligibility[is.na(Loan_Amount_Term), Loan_Amount_Term:=loanamountterm.mean]

# feature engineering (create LoanMonthly and Income columns)
loanamount.mean <- as.integer(mean(home.eligibility$LoanAmount, na.rm = TRUE))
loanamountterm.mean <- as.integer(mean(home.eligibility$Loan_Amount_Term, na.rm=TRUE))
home.eligibility[is.na(LoanAmount), LoanAmount:=loanamount.mean]
home.eligibility[is.na(Loan_Amount_Term), Loan_Amount_Term:=loanamountterm.mean]

# combine income
home.eligibility[,Income:=as.integer(ApplicantIncome+CoapplicantIncome)]
# create new loan amount
home.eligibility[,LoanMonthly:=LoanAmount/Loan_Amount_Term*30]
```

Factor Categorical columns
```{r}
cols <- c("Gender", "Married", "Dependents", "Education", "Self_Employed", "Property_Area", "Loan_Status")
home.eligibility[,cols] <- lapply(home.eligibility[,cols, with=FALSE], factor)
```

Data exploration revealed nearly 2:1 Y vs N, need to balance dataset
```{r}
# undersampling to balance the Y and N
home.eligibility[Loan_Status=="N",.N]
yes.id <- sample(home.eligibility[Loan_Status=="Y"]$Loan_ID, size=192)
home.eligibility.balanced <- home.eligibility[Loan_ID %in% yes.id | Loan_Status=="N"]

#verify
summary(home.eligibility.balanced)
```
Save file
```{r}
fwrite(home.eligibility.balanced, file="home_eligibility_prediction.csv")
```
