Try executing this chunk by clicking the *Run* button within the chunk or by placing your cursor inside it and pressing *Ctrl+Shift+Enter*. 

Create a new chunk by clicking the *Insert Chunk* button on the toolbar or by pressing *Ctrl+Alt+I*.

When you save the notebook, an HTML file containing the code and output will be saved alongside it (click the *Preview* button or press *Ctrl+Shift+K* to preview the HTML file).

Home Loan Data Set - Loan Status as label


```{r}
library(data.table)
library(ggplot2)
home.dt <- fread("home_train.csv")
#class(home.dt)
summary(home.dt)
sum(is.na(home.dt)) #86 missing values
apply(home.dt,2,function(x) any(is.na(x)))
#Id which variables have the most missing values and where
home.dt[is.na(LoanAmount),.N]

```

Columns - LoanAmount, Credit_History,Loan_Amount_Term having NA values but null values ignored

```{r}
#Change read 
home2.dt <- fread('home_train.csv', na.strings = c("NA", "missing", "N/A", -99, "", "m", "M", "na", "."))
sum(is.na(home2.dt)) #86 missing values
apply(home2.dt,2,function(x) any(is.na(x))) #Missing values show up
#is.na(home2.dt$Married)
summary(home2.dt$LoanAmount)
```
1. Any value in dropping the customer_ID? Because eventually want to identify these customers 
Clean vital few stats like: Gender: 
```{r}
summary(home2.dt$Gender)
# Value = Missing may be people who don't identify as male/female
# Create a new column complete_gender. Temporary recode NA as missing in complete_gender column.
#Convert to factor
home2.dt[Gender == 'Male',  complete_gender:= 'male']
home2.dt[Gender == 'Female', complete_gender:= 'female']
home2.dt[is.na(Gender), complete_gender:='missing']
home2.dt$Gender <- factor(home2.dt$complete_gender, levels = c('missing', 'male', 'female')) #encode as a factor
summary(home2.dt$Gender)
```
Clean Married as factor with 1 or 0, nominal variable:
```{r}
summary(home2.dt$Married)
home2.dt[Married == 'Yes',  complete_married:= '1']
home2.dt[Married == 'No', complete_married:= '0']
home2.dt[is.na(Married), complete_married:='2']
home2.dt$Married <- factor(home2.dt$complete_married, levels = c('0', '1', '2'))
summary(home2.dt$Married)
```
Clean Education
```{r}
summary(home2.dt$Education)
home2.dt[Education == 'Graduate',  complete_edu:= '1']
home2.dt[Education== 'Not Graduate', complete_edu:= '0']
home2.dt[is.na(Education), complete_edu:='2']
home2.dt$Education <- factor(home2.dt$complete_edu, levels = c('0', '1', '2'))
summary(home2.dt$Education)
```
Clean Self-Employed
```{r}
summary(home2.dt$Self_Employed)
home2.dt[Self_Employed == 'No',  biz_owner:= '0']
home2.dt[Self_Employed == 'Yes', biz_owner:= '1']
home2.dt[is.na(Self_Employed), biz_owner:='2']
home2.dt$Self_Employed <- factor(home2.dt$biz_owner, levels = c('0', '1', '2'))
summary(home2.dt$Self_Employed)
```
Clean Property_Area - could explore making these ordinal var? Unlikely to be true homeless people
```{r}
summary(home2.dt$Property_Area)
home2.dt[Property_Area == 'Urban',  prop:= 'U']
home2.dt[Property_Area == 'Semiurban', prop:= 'SU']
home2.dt[Property_Area == 'Rural', prop:= 'R']
home2.dt$Property_Area <- factor(home2.dt$prop, levels = c('U', 'SU', 'R','M'))
summary(home2.dt$Property_Area)
```
Clean Loan_Status 

```{r}
home2.dt$Loan_Status <- as.factor(home2.dt$Loan_Status)
```

Check status
2. Dummy variables for the categorical variables? 
```{r}
summary(home2.dt)
```

Clean continuous variables: LoanAmount, Loan_Amount_Term, Credit_History to then do distribution plots
Total of 22 + 14 + 50 missing cont variables => substantial? should find a way to deal with them? 

```{r}
ggplot(data = home2.dt[Loan_Status == "Y"], aes(x =ApplicantIncome , y = LoanAmount)) +
  geom_point() +
  #facet_grid(. ~ month) +
  labs(title = "Applicant Income vs Loan Amount")
```
^ Looks interesting, hence explore any other linear r/s never considered using a correlation matrix? 

Other data sets used log transform to drop outliers 

```{r}
pairs(~ Gender+Married+Education+Self_Employed+ApplicantIncome+CoapplicantIncome+LoanAmount+Loan_Amount_Term+Credit_History+Property_Area, panel=panel.smooth, data = home2.dt)

```
```{r}
library(corrplot)
M<- cor(home2.dt[,c('ApplicantIncome','CoapplicantIncome','LoanAmount','Loan_Amount_Term')],use='complete.obs')
corrplot(M, method="color",addCoef.col = "black",tl.col = "black")
```


```{r}
# Density Plot to see distribution of continuous varable.
ggplot(home2.dt,aes(ApplicantIncome)) + geom_density()
which.max(density(home2.dt$ApplicantIncome)$y)
density(home2.dt$ApplicantIncome)$x[29]
cor(home2.dt$CoapplicantIncome,home2.dt$ApplicantIncome)
summary(home2.dt)
```

